A framework for adaptive width control of dense contour-parallel toolpaths in additive manufacturing

### 2019-08-30 12_conclusions_future_work.tex
Tim: nominal => preferred
'nominal size' is the size something is supposed to be on paper; that wouldn't work for a sentence like
'The evenly distributed scheme can be conceptualized as calculating the total discrepancy $ between the actual feature diameter $ and the total preferred width  w^*$'

We presented the inward distributed beading scheme.
It reduces the amount of beads with a width deviating extremely from the nominal bead width, and thus it is expected to limit the impact of adaptive bead width on mechanical properties of the resulting print.
Furthermore, by distributing the deviation in bead width over several beads near the center of the shape, the outer toolpath is affected less by the deviation, meaning that the dimensional accuracy in the horizontal direction is affected less by inaccuracies in the mechanical system which realizes adaptive bead width.

>>>>

We presented the inward distributed beading scheme.
It reduces the amount of beads with a width deviating extremely from the preferred bead width, and thus it is expected to limit the impact of adaptive bead width on mechanical properties of the resulting print.
Furthermore, by distributing the deviation in bead width over several beads near the center of the shape, the outer toolpath is affected less by the deviation, meaning that the dimensional accuracy in the horizontal direction is affected less by inaccuracies in the mechanical system which realizes adaptive bead width.

### 2019-08-30 1_intro.tex
Tim: nominal => preferred
'nominal size' is the size something is supposed to be on paper; that wouldn't work for a sentence like
'The evenly distributed scheme can be conceptualized as calculating the total discrepancy $ between the actual feature diameter $ and the total preferred width  w^*$'

\section{Introduction}
3D printing enables the fabrication of complex geometry under few design constraints compared to conventional fabrication techniques.
Recent developments have seen a rapid growth in both the use and capabilities of desktop 3D printing systems.
%The rapid spread of 3D printing through different industries and types of application calls for the possibility to manufacture a wide range of geometries while guaranteeing mechanical properties of the resulting parts.
Fused Deposition Modeling (FDM) is one of the most common 3D printing techniques.
It is widely used because of the versatility in the types of plastics which can be used and the relatively low running costs.
FDM printers are used, for example, in showcasing scale models of buildings, casings for electronics, prototypes for blow molded parts, jigs and fixtures.
Recent research developments have investigated manufacturing complex volumetric structures such as microstructures~\cite{bates2018compressive,Liu2019a}, topology optimized structures~\cite{Zong2019,Wu2019a,Cheng2019} and regular honeycomb structures~\cite{Al-Ketan2018,Maskery2018}.
Many of these applications require 3D models with detailed features within the order of magnitude of the printing resolution, which restrains the field of the process planning algorithms.

FDM printers extrude semi-continuous beads of molten plastic through a nozzle, which moves along a planned toolpath within each layer of a 3D object.
A common strategy to accurately manufacture a given 3D model is to extrude a contour-following print path,
because the position and shape of the toolpath can be controlled relatively accurately.
Filling up a shape using parallel straight lines would expose defects of the size of the nozzle ($\pm 10^2 \mu$), which is generally an order of magnitude larger than the resolution of the positioning system ($\pm 10^1 \mu$).
Contour parallel extrusion therefore leads to a less bumpy outline shape than direction parallel extrusion does.

The simple technique for generating the shape filling toolpaths of a layer consists in performing uniform inward offsets with the radius of the the nozzle from the outline shape.
However, for geometrical features which are not an exact multiple of the nozzle size this method produces areas where an extrusion bead is placed twice: \emph{overfill} areas; and other areas which are not filled at all: \emph{underfill} areas.
See \cref{intro_wedge}(top).
Overfills cause a buildup of pressure in the mechanical extrusion system, which can result in bulges or even a full print failure.
Underfills on the other hand, can cause a drastic decrease in the part stiffness or even for small features not to be printed at all.
These problems are exacerbated for models with layer outlines with small features, because the over- and underfill areas are relatively large compared to the whole part.
In order to avoid over- and underfills in regions which are not as wide as an exact multiple of the nozzle size we need algorithms which produce toolpaths which employ adaptive bead width.
Some such toolpath generation techniques have already been developed.

@@

There is a large body of literature on relating process parameters such as bead width to various mechanical properties of the end part.
The bead width has been associated with total print time, the roughness of top surfaces, printing resolution and the strength of a part, through the contact area it makes with neighboring beads~\cite{N.Turner2014,ahn2002anisotropic}.
In order to accurately fabricate a part restricted to prescribed mechanical properties, we need to generate toolpaths using the corresponding process parameters as close as possible.
We therefore develop a framework which can be used to generate toolpaths which not only avoid over- and underfill, but also keep the bead width closer to the nominal bead width compared to existing techniques.
See \cref{intro_wedge}(bottom).

%\subsection{Contributions}
A polygonal shape with a width of a fractional multiple of the nominal bead width will always result in the bead widths deviating from the nominal bead width by the fractional part in total,
because the total of the widths of the beads must add up to the fractional width of the part in order to avoid under- or overfill.
We therefore propose a framework which makes it possible to distribute the fractional discrepancy over several beads, so that the beads are generally closer to the nominal width.
Our contributions are as follows:
\begin{itemize}
\item A geometric framework for generating densely filling contour-parallel toolpaths employing adaptive width, according to any function which maps a model diameter to a number of beads and their widths.
\item A specific beading scheme function for this framework for:
\begin{itemize}
\item Accurately filling a 2D outline with extrusion beads
\item Reducing the overfill and underfill areas compared to the uniform width technique
\item Reducing the amount of beads with greatly deviating width compared to existing literature
\item Promoting smooth toolpaths which are close to the nominal width toward the outline of the shape.
\end{itemize}
\end{itemize}

>>>>

\section{Introduction}
3D printing enables the fabrication of complex geometry under few design constraints compared to conventional fabrication techniques.
Recent developments have seen a rapid growth in both the use and capabilities of desktop 3D printing systems.
%The rapid spread of 3D printing through different industries and types of application calls for the possibility to manufacture a wide range of geometries while guaranteeing mechanical properties of the resulting parts.
Fused Deposition Modeling (FDM) is one of the most common 3D printing techniques.
It is widely used because of the versatility in the types of plastics which can be used and the relatively low running costs.
FDM printers are used, for example, in showcasing scale models of buildings, casings for electronics, prototypes for blow molded parts, jigs and fixtures.
Recent research developments have investigated manufacturing complex volumetric structures such as microstructures~\cite{bates2018compressive,Liu2019a,Al-Ketan2018,Maskery2018} and topology optimized structures~\cite{Zong2019,Wu2019a,Cheng2019}.
Many of these applications involve 3D models with detailed features within the order of magnitude of the printing resolution, which restrains the field of the process planning algorithms.

FDM printers extrude semi-continuous beads of molten plastic through a nozzle, which moves along a planned toolpath within each layer of a 3D object.
A common strategy to accurately manufacture a given 3D model is to extrude along a contour-following path,
because the position and shape of the toolpath can be controlled relatively accurately.
Filling up a shape using parallel straight lines would expose defects of the size of the nozzle ($\pm 10^2 \mu$), which is generally an order of magnitude larger than the resolution of the positioning system ($\pm 10^1 \mu$).
Contour parallel extrusion therefore leads to a less bumpy outline shape than direction parallel extrusion does.

The simple technique for generating the dense contour-parallel toolpaths of a layer consists in performing uniform inward offsets with the radius of the nozzle from the outline shape.
However, for geometrical features which are not an exact multiple of the nozzle size this method produces areas where an extrusion bead is placed twice: \emph{overfill} areas; and areas which are not filled at all: \emph{underfill} areas.
See \cref{intro_wedge}(top).
Overfills cause a buildup of pressure in the mechanical extrusion system, which can result in bulges or even a full print failure.
Underfills on the other hand, can cause a drastic decrease in the part stiffness or even for small features not to be printed at all.
These problems are exacerbated for models with layer outlines with small features, because the over- and underfill areas are relatively large compared to the whole part.

One promising direction to avoid over- and underfills is to employ toolpath with adaptive width. 
This strategy has been mostly investigated for metal additive manufacturing using wire and arc~\cite{Ding2014}, where a wide range of widths could be realized by the hardware. 
For instance, \citeauthor{Ding2016a} developed a toolpath with an (estimated) width variation by a factor of up to $3$.
However, for FDM printing, it is challenging to realize varying width by extruding through a nozzle with a fixed size (the preferred width), either by varying the movement speed or by dynamically adjusting the volumetric flow rate. 
Therefore, a much narrower range of widths is desirable for robust FDM fabrication.
On the other hand, controlling the variation in width is beneficial for limiting the variation in mechanical properties of the resulting material fragments. 
% The bead width has been associated with total print time, the roughness of top surfaces, printing resolution and the strength of a part, through the contact area it makes with neighboring beads~\cite{N.Turner2014,ahn2002anisotropic}.

In this paper we propose a framework for planning toolpath with adaptive width for minimizing over- and underfill. 
We show that this framework supports multiple schemes on deciding the bead locations and extrusion widths. 
In particular for FDM printing, we propose a novel scheme which reduces the amount of over- and underfill while the extrusion widths deviate less from nozzle size compared to existing approaches (which are mostly designed for wire and arc additive manufacturing). 
% In order to avoid over- and underfills in regions which are not as wide as an exact multiple of the nozzle size we need algorithms which produce toolpaths which employ adaptive bead width.
% Some such toolpath generation techniques have already been developed.

@@

% There is a large body of literature on relating process parameters such as bead width to various mechanical properties of the end part.
% The bead width has been associated with total print time, the roughness of top surfaces, printing resolution and the strength of a part, through the contact area it makes with neighboring beads~\cite{N.Turner2014,ahn2002anisotropic}.
% In order to accurately fabricate a part restricted to prescribed mechanical properties, we need to generate toolpaths using the corresponding process parameters as close as possible.
% We therefore develop a framework which can be used to generate toolpaths which not only avoid over- and underfill, but also keep the bead width closer to the preferred bead width compared to existing techniques.
% See \cref{intro_wedge}(bottom).

%\subsection{Contributions}
% A polygonal shape with a width of a fractional multiple of the preferred bead width will always result in the bead widths deviating from the preferred bead width by the fractional part in total,
% because the total of the widths of the beads must add up to the fractional width of the part in order to avoid under- or overfill.
% We therefore propose a framework which makes it possible to distribute the fractional discrepancy over several beads, so that the beads are generally closer to the preferred width.
Our contributions are as follows:
\begin{itemize}
\item A geometric framework for generating densely filling contour-parallel toolpaths employing adaptive width, according to any beading scheme which maps a feature diameter to a number of beads and their widths.
\item A specific beading scheme for FDM printing which reduces the amount of beads with greatly deviating width, and which promotes smooth toolpaths that are close to the preferred width toward the outline of the shape.
% \begin{itemize}
% \item Accurately filling a 2D outline with extrusion beads
% \item Reducing the overfill and underfill areas compared to the uniform width technique
% \item Reducing the amount of beads with greatly deviating width compared to existing literature
% \item Promoting smooth toolpaths which are close to the preferred width toward the outline of the shape.
% \end{itemize}
\end{itemize}

### 2019-08-30 8_discussion.tex
Small changes by Jun and Tim
Tim: nominal => preferred
'nominal size' is the size something is supposed to be on paper; that wouldn't work for a sentence like
'The evenly distributed scheme can be conceptualized as calculating the total discrepancy $ between the actual feature diameter $ and the total preferred width  w^*$'

The constant bead count scheme effectively deals with underfills, but generates orders of magnitude more overfills compared to the other schemes. 
Also, the scheme comes at the cost of greatly varying bead widths and an average bead width that is not close to the nominal bead width.
Note that most overfill areas occur near regions of alternating bead width. 
For an input outline shape which contains both very small and very large features, the the constant bead count scheme produces bead widths which can fall outside of the range of manufacturable bead widths.
Moreover the centrality marking is not robust against small perturbations in the outline; adding a small chamfer in a corner causes the unmarked ST to be very small at that location, which results in tiny bead widths.

@@

However, according to \cref{widthHistogram} the centered scheme exhibits a wider range of bead widths than the distributed schemes:
the standard deviation of the bead widths in the centered scheme is approximately \SI{39}{\micro\meter}, while that of the distributed schemes is approximately \SI{14}{\micro\meter}.
We can therefore state that the distributed schemes result in bead widths closer to the nominal widths compared to the centered scheme.
This is desirable for the manufacturability of the beads and can therefore have a positive effect on the mechanical properties and surface quality of the 3D prints.

The inward distributed scheme is introduced with the intention to limit the number of beads deviating from the nominal bead width and to limit the number of angled extrusion segments in transitioning regions.
This expected effect is difficult to be confirmed when comparing \cref{TEST_Distributed_accuracy} with~\ref{TEST_InwardDistributed_accuracy}, because there are no large solid areas in that specific shape.

However, when comparing the bead width of the outer toolpaths, \cref{distributed_comparison} and  \cref{widthIndexedHistogram}, we can clearly see that the these toolpaths have the nominal bead width more often, which means that errors from inaccuracies due to deviation from the nominal bead width are concentrated near the center of the shape, leaving the dimensional accuracy of the outline shape less affected, if at all. 
Furthermore, we find that compared to evenly distributed, the inward distributed scheme produces smoother toolpaths overall (\cref{smoothness}), and with less jaggedness on the contour toolpaths (\cref{smoothness}). 
Therefore, when using the inward distributed scheme the surface of the 3D print is less affected by both width variation and jagged toolpaths, while overfill and underfill are kept at a low percentage.

@@

\subsection{Limitations}
Note that the current industry standard in FDM printing employs little to no bead width variation.
Properly performing bead width variation calls for adaptations and developments in printers and firmware.
In the beading schemes we set a transition length of $t(n) = w^*$.
That will demand changes in cross-sectional area of the bead up to \SI{200}{\percent} over a small distance that is comparable to the nozzle size, which is challenging for some hardware systems.
Varying the movement speed can be utilised to change the cross-sectional area, but this approach is limited, since the movement speed is constrained by acceleration considerations near bends in the toolpath.
Our schemes require a more accurate control of the volumetric flow rate in \si{\milli\meter\cubed\per\second}.
Using a filament feeder directly mounted on the print head (a.k.a. direct drive) outperforms Bowden style FDM printers in terms of dynamic flow changes,
but still requires some control system in order to accurately change the volumetric flow rate such as pressure advance algorithms~\cite{tronvoll2019investigating}.
Yet inaccuracies in direct drive systems employing advance algorithms might arise due to the changes in back-pressure required by changing bead size.
We expect that developments in printing hardware and firmware will address these challenges in the future.

>>>>

The constant bead count scheme effectively deals with underfills, but generates orders of magnitude more overfills compared to the other schemes. 
Also, the scheme comes at the cost of greatly varying bead widths and an average bead width that is not close to the preferred bead width.
Note that most overfill areas occur near regions of alternating bead width. 
For an input outline shape which contains both very small and very large features, the the constant bead count scheme produces bead widths which can fall outside of the range of manufacturable bead widths.
Moreover the centrality marking is not robust against small perturbations in the outline; adding a small chamfer in a corner causes the unmarked ST to be very small at that location, which results in tiny bead widths.

@@

However, according to \cref{widthHistogram} the centered scheme exhibits a wider range of bead widths than the distributed schemes:
the standard deviation of the bead widths in the centered scheme is approximately \SI{39}{\micro\meter}, while that of the distributed schemes is approximately \SI{14}{\micro\meter}.
We can therefore state that the distributed schemes result in bead widths closer to the preferred widths compared to the centered scheme.
This is desirable for the manufacturability of the beads and can therefore have a positive effect on the mechanical properties and surface quality of the 3D prints.

The inward distributed scheme is introduced with the intention to limit the number of beads deviating from the preferred bead width and to limit the number of angled extrusion segments in transitioning regions.
This expected effect is difficult to be confirmed when comparing \cref{TEST_Distributed_accuracy} with~\ref{TEST_InwardDistributed_accuracy}, because there are no large solid areas in that specific shape.

However, when comparing the bead width of the outer toolpaths, \cref{distributed_comparison} and  \cref{widthIndexedHistogram}, we can clearly see that the these toolpaths have the preferred bead width more often, which means that errors from inaccuracies due to deviation from the preferred bead width are concentrated near the center of the shape, leaving the dimensional accuracy of the outline shape less affected, if at all. 
Furthermore, we find that compared to evenly distributed, the inward distributed scheme produces smoother toolpaths overall (\cref{smoothness}), and with less jaggedness on the contour toolpaths (\cref{smoothness}). 
Therefore, when using the inward distributed scheme the surface of the 3D print is less affected by both width variation and jagged toolpaths, while overfill and underfill are kept at a low percentage.

@@

\subsection{Discussion on implications}
Note that the current industry standard in FDM printing employs little to no bead width variation.
Properly performing bead width variation calls for adaptations and developments in printers and firmware.
In the beading schemes we set a transition length of $t(n) = w^*$.
That will demand changes in cross-sectional area of the bead up to \SI{200}{\percent} over a small distance that is comparable to the nozzle size, which is challenging for some hardware systems.
Varying the movement speed can be utilised to change the cross-sectional area, but this approach is limited, since the movement speed is constrained by acceleration considerations near bends in the toolpath.
Our schemes require a more accurate control of the volumetric flow rate in \si{\milli\meter\cubed\per\second}.
Using a filament feeder directly mounted on the print head (a.k.a. direct drive) outperforms Bowden style FDM printers in terms of dynamic flow changes,
but still requires some control system in order to accurately change the volumetric flow rate such as pressure advance algorithms~\cite{tronvoll2019investigating}.
Yet inaccuracies in direct drive systems employing advance algorithms might arise due to the changes in back-pressure required by changing bead size.
We expect that developments in printing hardware and firmware will address these challenges in the future.
